What We Eval in the Shadows

### 2021-04-12 analysis/environments.Rmd
Explain the various environments wrapped with new

```{r}
caller_envs <- env_calls %>% filter(!is.na(hierarchy), is.na(specific_env) || specific_env == "_callee", is.na(new_env))

@@

```{r}
# TODO: high-level per caller frame number...
```

>>>>

```{r}
caller_envs <- env_calls %>% filter(!is.na(hierarchy), is.na(specific_env) | specific_env == "_callee", is.na(new_env))

@@

```{r}
caller_env_sites <- caller_envs %>% group_by(hierarchy) %>% 
  summarise(n = n_distinct(src_ref)) %>% 
  arrange(desc(n)) %>% 
   mutate(pc = ratio(n, nb_caller_env_sites), cpc = cumsum(pc))

nrows <- 5
r_vec("caller env hierarchy name", nrows, caller_env_sites$hierarchy)
r_vec("caller env hierarchy sites", nrows, caller_env_sites$n)
r_vec("caller env hierarchy site percent", nrows, caller_env_sites$pc)

nb_far_away_caller_sites <- caller_envs %>% filter(hierarchy > 3 | is.na(hierarchy)) %>% pull(src_ref) %>% n_distinct()

overview_table(
  r("nb far away caller sites", nb_far_away_caller_sites),
  r("nb far away caller site percent", ratio(nb_far_away_caller_sites, nb_caller_env_sites))
)
```

### 2021-04-12 paper/main.tex
Explain the various environments wrapped with new

\begin{itemize}
	\item function environments: the frame of the calling function or one of the parent frames (obtained using \c{parent.frame()}, \c{sys.frame()} or with an integer argument for \c{envir} for instance);
	\item list environments: a list, a data frame, a tibble, or \c{NULL} (seen as the empty list), passed to \eval;
	\item the global environment;
	\item package and namespace environments;
	\item the empty environment (created with \c{emptyenv()})
\end{itemize}

@@

The calls happening in a function environment are mostly happening in the environment of the function calling \eval, as shown in Table%TODO

%TODO: add here table of possible frame numbers

Most of the calls happen in the environment surrounding \eval. This corresponds mostly to the default \c{envir} argument of \eval; indeed, \packageDefautEnvirSitesPercent of the sites have that default argument. A lot of sites also use \c{parent.frame()} to access their parent frame. It corresponds to environment class \c{1} but also to \c{global} when the function is called at the top level. Explicitly and directly referring to the global environment using \c{globalenv()} or \c{.GlobalEnv} is rare, as it happens only in \packageNbExplicitGlobalSites sites. Accessing further-away frames happens much more seldom.

%TODO: add here table of possible environment categories wrapped in new

\begin{table}[h]
    \centering 
    \begin{tabular}{|c|c|c|}
        \hline
        Category & \#sites & \%sites \\
        \hline
        \packageNewEnvCategoryNamea & \packageNewEnvCategorySitesa & \packageNewEnvCategorySitePercenta \\
        \hline
        \packageNewEnvCategoryNameb & \packageNewEnvCategorySitesb &  \packageNewEnvCategorySitePercentb\\
        \hline
        \packageNewEnvCategoryNamec & \packageNewEnvCategorySitesc & \packageNewEnvCategorySitePercentc \\
        \hline
        \packageNewEnvCategoryNamed & \packageNewEnvCategorySitesd & \packageNewEnvCategorySitePercentd \\
        \hline
    \end{tabular}
\caption{Classification of environments wrapped by one of several \c{new}, in number of sites. Some sites can be passed more than one kind of environments. }
\end{table}

% Up-to-date
\begin{table}[htbp]{ \centering
        \begin{tabular}{|c|c|}\hline
            Environment Class & Sites\\\hline
            \packageEnvClassNamesa &  \packageEnvClassSitePercenta\\
            \packageEnvClassNamesb &  \packageEnvClassSitePercentb\\
            \packageEnvClassNamesc &  \packageEnvClassSitePercentc\\
            \packageEnvClassNamesd &  \packageEnvClassSitePercentd\\
            \packageEnvClassNamese &  \packageEnvClassSitePercente\\
            \packageEnvClassNamesf &  \packageEnvClassSitePercentf\\
            \packageEnvClassNamesg &  \packageEnvClassSitePercentg\\
            \packageEnvClassNamesh &  \packageEnvClassSitePercenth\\
            \hline
    \end{tabular}} \label{tab:environments}
    \caption{Environments in terms of sites. One site can be passed different kinds of environments.}
\end{table}

>>>>

\begin{itemize}
	\item function environments: the frame of the calling function or one of the parent frames (obtained using \c{parent.frame()}, \c{sys.frame()} or with an integer argument for \c{envir} for instance);
	\item list environments: a list, a data frame, a tibble, or \c{NULL} (seen as the empty list), passed to \eval;
	\item the global environment;
	\item package and namespace environments;
	\item the empty environment (created with \c{emptyenv()})
    \item new environments, created with \c{new.env}, \c{list2env} or \c{as.environment}.
\end{itemize}

@@

Table~\ref{tab:framenumber} shows the function environments classified according to their position in the call stack: $0$ is the current function, $1$ the parent function and so on.

The calls happening in a function environment are mostly happening in the environment of the function calling \eval. This corresponds mostly to the default \c{envir} argument of \eval; indeed, \packageDefautEnvirSitesPercent of the sites have that default argument. A lot of sites also use \c{parent.frame()} to access their parent frame. It corresponds to environment class \c{1} but also to \c{global} when the function is called at the top level. Explicitly and directly referring to the global environment using \c{globalenv()} or \c{.GlobalEnv} is rare, as it happens only in \packageNbExplicitGlobalSites sites. Accessing further-away frames happens much more seldom.

\begin{table}[h]
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        Frame number & \#sites & \%sites \\
        \hline
        \packageCallerEnvHierarchyNamea & \packageCallerEnvHierarchySitesaRnd & \packageCallerEnvHierarchySitePercenta \\
        \packageCallerEnvHierarchyNameb& \packageCallerEnvHierarchySitesbRnd & \packageCallerEnvHierarchySitePercentb \\
        \packageCallerEnvHierarchyNamec& \packageCallerEnvHierarchySitescRnd & \packageCallerEnvHierarchySitePercentc  \\
        > 3& \packageNbFarAwayCallerSites &  \packageNbFarAwayCallerSitePercent \\
        \hline
    \end{tabular} \label{tab:framenumber}
\caption{Frame number for the function environments. One site can possibly be passed different kinds of environments.}
\end{table}

A new environment creation, a parent environment must be specified. In Table~\ref{tab:newenvs}, we show what environments the new environment inherits from ultimately.\footnote{It means we directly show the first parent that is different from a new environment.} New environments mostly inherits from function environment: that is actually the default for \c{new.env}, with argument \c{parent = parent.frame()}. A large number of new environments are also wrapping package environment or namespaces. This is a widespread pattern to store variables to use at the package level. When a new environment wraps the empty environment, it can be handily used as a hash table.\footnote{\c{new.env} builds an environment that stores the variable name with a hash table by default. }
% Speak also about new and global?

\begin{table}[h]
    \centering 
    \begin{tabular}{|c|c|c|}
        \hline
        Category & \#sites & \%sites \\
        \hline
        \packageNewEnvCategoryNamea & \packageNewEnvCategorySitesa & \packageNewEnvCategorySitePercenta \\
        \packageNewEnvCategoryNameb & \packageNewEnvCategorySitesb &  \packageNewEnvCategorySitePercentb\\
        \packageNewEnvCategoryNamec & \packageNewEnvCategorySitesc & \packageNewEnvCategorySitePercentc \\
        \packageNewEnvCategoryNamed & \packageNewEnvCategorySitesd & \packageNewEnvCategorySitePercentd \\
        \hline
    \end{tabular} \label{tab:newenvs}
\caption{Classification of environments wrapped by one or several \c{new}, in number of sites. Some sites can be passed more than one kind of environments. }
\end{table}

% Up-to-date
% Probably to remove
\begin{table}[htbp]{ \centering
        \begin{tabular}{|c|c|}\hline
            Environment Class & Sites\\\hline
            \packageEnvClassNamesa &  \packageEnvClassSitePercenta\\
            \packageEnvClassNamesb &  \packageEnvClassSitePercentb\\
            \packageEnvClassNamesc &  \packageEnvClassSitePercentc\\
            \packageEnvClassNamesd &  \packageEnvClassSitePercentd\\
            \packageEnvClassNamese &  \packageEnvClassSitePercente\\
            \packageEnvClassNamesf &  \packageEnvClassSitePercentf\\
            \packageEnvClassNamesg &  \packageEnvClassSitePercentg\\
            \packageEnvClassNamesh &  \packageEnvClassSitePercenth\\
            \hline
    \end{tabular}} \label{tab:environments}
    \caption{Environments in terms of sites. One site can be passed different kinds of environments.}
\end{table}
